---
title: "CoDA_HandlingDegradedCells"
output: html_document
---

load library
```{r message=FALSE, warning=FALSE}
library(Seurat)
library(ggplot2)
library(ggpubr)
library(SingleCellExperiment)
library(scran)
library(SeuratData)
library(edgeR)
library(limma)
library(clusterProfiler)
library(data.table)
library(SummarizedExperiment)
library(muscat)
library(pROC)
library(igraph)
library(mclust)
library(cluster)
library(SCINA)
library(caret)
library(robCompositions)
library(easyCODA)
library(MAST)
library(AUCell)
library(HGNChelper)
library(SeuratDisk)
library(scmap)
library(monocle3)
library(slingshot)
library(grDevices)
library(RColorBrewer)
library(randomForest)
library(caTools)
library(mltools)
library(reldist)
library(aricode)
library(TSCAN)
library(reshape2)
library(pheatmap)
library(philentropy)
library(rawr)
library(splatter)
library(combinat)
library(Hmisc)
library(plotly)
library(ISnorm)
library(dbscan)
library(ca)
library(factoextra)
```

functions
```{r message=FALSE, warning=FALSE}
get_earliest_principal_node <- function(cds, time_bin="C8"){
  cell_ids <- which(colData(cds)[,"cluster"] == time_bin)
  
  closest_vertex <-
  cds@principal_graph_aux[["UMAP"]]$pr_graph_cell_proj_closest_vertex
  closest_vertex <- as.matrix(closest_vertex[colnames(cds), ])
  root_pr_nodes <-
  igraph::V(principal_graph(cds)[["UMAP"]])$name[as.numeric(names
  (which.max(table(closest_vertex[cell_ids,]))))]
  
  root_pr_nodes
}

getCluster <- function(x,meta,reso=0.01,npc=10,cent=4,scale_data=FALSE){
    tmp <- CreateSeuratObject(counts=x,meta.data=meta,min.cells=0,min.features=0)
    tmp <- FindVariableFeatures(tmp,nfeatures=3000,verbose=FALSE)
    if(scale_data){
        tmp <- ScaleData(tmp,verbose=FALSE)
    }else{
        tmp[["RNA"]]@scale.data <- as.matrix(tmp[["RNA"]]@counts[VariableFeatures(tmp),])
    }
    tmp <- RunPCA(tmp,verbose=FALSE)
    tmp$kmeans <- kmeans(tmp@reductions$pca@cell.embeddings[,1:npc],centers=cent)$cluster
    tmp <- FindNeighbors(tmp,dims=1:npc,verbose=FALSE)
    tmp <- FindClusters(tmp,resolution=reso,verbose=FALSE)
    tmpgraph <- buildSNNGraph(tmp@reductions$pca@cell.embeddings[,1:npc],transposed=T,k=10,d=NA)
    res <- cluster_louvain(tmpgraph)$membership
    cc <- aggregate(tmp@reductions$pca@cell.embeddings[,1:npc],list(res),mean)
    cc <- as.matrix(cc[,-1])
    hclu <- hclust(dist(cc))
    clu <- cutree(hclu,cent)
    clu <- clu[res]      
    tmp$lv_clu <- clu
    #tmp <- tmp@meta.data
    tmp <- RunUMAP(tmp,dims=1:npc,verbose=FALSE)
    return(tmp)
}

evalClustering <- function(x,npc=10,method="lv_clu"){
    
    acc <- -mean(sapply(unique(x@meta.data[,method]),function(i){
        p <- table(x@meta.data[x@meta.data[,method]==i,"celltype"])/sum(x@meta.data[,method]==i)
        sum(p*log(p),na.rm=TRUE)
    }))

    pur <- -mean(sapply(unique(x$celltype),function(sct){
        p <- table(x@meta.data[x$celltype==sct,method])/sum(x$celltype==sct)
        sum(p*log(p),na.rm=TRUE)
    }))

    ARI <- adjustedRandIndex(x$celltype,x@meta.data[,method])

    d <- dist(x@reductions$pca@cell.embeddings[,1:npc])
    s <- silhouette(x@meta.data[,method],dist=d)
    medianSil <- median(s[,3])
    
    NMI_value <- NMI(x@meta.data[,method],x$celltype)
    #NMI_value <- NMI(x$lv_clu,x$celltype,variant="sum")
    
    df <- c(acc,pur,ARI,medianSil,NMI_value)
    names(df) <- c("acc","pur","ARI","medianSil","NMI")
    return(df)
}

evalClustering_twoGroup <- function(x,npc=10,method="lv_clu"){
    
    acc <- -mean(sapply(unique(x@meta.data[,method]),function(i){
        p <- table(x@meta.data[x@meta.data[,method]==i,"label"])/sum(x@meta.data[,method]==i)
        sum(p*log(p),na.rm=TRUE)
    }))

    pur <- -mean(sapply(unique(x$label),function(sct){
        p <- table(x@meta.data[x$label==sct,method])/sum(x$label==sct)
        sum(p*log(p),na.rm=TRUE)
    }))

    ARI <- adjustedRandIndex(x$label,x@meta.data[,method])

    d <- dist(x@reductions$pca@cell.embeddings[,1:npc])
    s <- silhouette(x@meta.data[,method],dist=d)
    medianSil <- median(s[,3])
    
    NMI_value <- NMI(x@meta.data[,method],x$label)
    #NMI_value <- NMI(x$lv_clu,x$celltype,variant="sum")
    
    df <- c(acc,pur,ARI,medianSil,NMI_value)
    names(df) <- c("acc","pur","ARI","medianSil","NMI")
    return(df)
}
```

# Handle degraded cells
## cellbench 
read and process
```{r}
bulk_dat <- readRDS("GSE86337_processed_count_average_replicates.rds")

##raw
dat <- fread("cellbench_sc_10x_5cl_genebycell.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
dat <- as.matrix(dat,rownames=1)

metadat <- sapply(colnames(dat),function(x) strsplit(x,"\\.")[[1]][2])
metadat <- data.frame(cell=colnames(dat),celltype=metadat)

bulk_dat <- bulk_dat[rownames(bulk_dat)%in%rownames(dat),]
dat <- dat[rownames(bulk_dat),]

#raw lognorm
dat <- CreateSeuratObject(dat,meta.data=metadat,min.cells=0,min.features=0)
dat <- NormalizeData(dat)
dat <- FindVariableFeatures(dat,nfeatures=3000)
dat <- ScaleData(dat)
dat <- RunPCA(dat)
dat <- RunUMAP(dat,dims=1:50)
```

### randomly assign top gene zeros for each cell
10% H1975 cells and 6000 genes (~60%)
```{r}
#rank expressed features in selected types
SelectedType <- rownames(dat@meta.data[dat@meta.data$celltype=="H1975",])
SelectedTypeMat <- as.matrix(dat@assays$RNA@counts)
SelectedTypeMat <- SelectedTypeMat[,SelectedType]
TopFeatures <- apply(SelectedTypeMat,1,sum)
names(TopFeatures) <- rownames(SelectedTypeMat)
TopFeatures <- names(TopFeatures[order(TopFeatures,decreasing=TRUE)])

#select 10% of cells and copy
set.seed(3)
Selected10Cells <- sample(x=SelectedType,size=0.1*length(SelectedType))
SelectedCellMat <- SelectedTypeMat[,Selected10Cells]

#assign zeros
RandomAssignZeros <- function(x,g,gname){
    names(x) <- gname
    SelectedGenes <- sample(g,size=6000)
    x[SelectedGenes] <- 0
    return(x)
}

ProcessedCellMat <- apply(SelectedCellMat,2,RandomAssignZeros,TopFeatures,rownames(SelectedCellMat))
colnames(ProcessedCellMat) <- paste0(colnames(ProcessedCellMat),".copy")

#combine
tmp <- cbind(as.matrix(dat@assays$RNA@counts),ProcessedCellMat)
tmpmeta <- data.frame(cell=colnames(ProcessedCellMat),celltype=rep("H1975.zero",ncol(ProcessedCellMat)))
rownames(tmpmeta) <- tmpmeta$cell
tmpmeta <- rbind(dat@meta.data[,c("cell","celltype")],tmpmeta)

#plot
tmp <- CreateSeuratObject(tmp,meta.data=tmpmeta,min.cells=0,min.features=0)

tmp <- NormalizeData(tmp)
tmp <- FindVariableFeatures(tmp,nfeatures=3000)
tmp <- ScaleData(tmp)
tmp <- RunPCA(tmp)
tmp <- RunUMAP(tmp,dims=1:50)

#sct
tmp_sct <- CreateSeuratObject(counts=tmp@assays$RNA@counts,meta.data=tmpmeta,min.cells=0,min.features=0)
tmp_sct <- SCTransform(tmp_sct,vst.flavor="v2",verbose=FALSE,min_cells=0)
tmp_sct <- RunPCA(tmp_sct)
tmp_sct <- RunUMAP(tmp_sct,dims=1:50)
```

CoDA
```{r}
#add pseudocount s/gm to raw count
tmp1 <- as.matrix(tmp@assays$RNA@counts)
gm_s <- exp(mean(log(apply(tmp1,2,sum))))
tmp1 <- apply(tmp1,2,function(x) x+(sum(x)/gm_s) )
tmp1 <- t(CLOSE(t(tmp1)))

#CLR
raw_coda_clr <- apply(tmp1,2,function(x) x/exp(mean(log(x))) )
raw_coda_clr <- log2(raw_coda_clr)

# HKGLR
HKGenes <- which(rownames(tmp1)%in%c("ACTB","ARBP","GAPDH","HPRT","SDHA","UBC","YWHAZ"))
raw_coda_ilr <- apply(tmp1,2,function(x) x/exp(mean(log(x[HKGenes]))) )
raw_coda_ilr <- log2(raw_coda_ilr)

##ILR
cb_ILR <- as.matrix(t(coda.base::coordinates(t(tmp1))))
cb_ILR <- cb_ILR[,colnames(raw_coda_clr)]

#Seurat process
processSeurat <- function(x,metadat){
    res <- CreateSeuratObject(x,meta.data=metadat,min.cells=0,min.features=0)
    res <- FindVariableFeatures(res,nfeatures=3000)
    res <- ScaleData(res)
    res <- RunPCA(res)
    res <- RunUMAP(res,dims=1:50)
    return(res)
}

raw_coda_clr <- processSeurat(raw_coda_clr,tmp@meta.data)
raw_coda_ilr <- processSeurat(raw_coda_ilr,tmp@meta.data)
cb_ILR <- processSeurat(cb_ILR,tmp@meta.data)

alldat <- list(Raw_LogNorm=tmp,Raw_SCT=tmp_sct,SGM_CLR=raw_coda_clr,SGM_ILR=cb_ILR,SGM_HKGLR=raw_coda_ilr)
alldat <- alldat[c("Raw_LogNorm","Raw_SCT","SGM_CLR","SGM_ILR","SGM_HKGLR")]
alldat_CellBench <- alldat
```

plots
umap
```{r fig.height=7.5, fig.width=8.5}
p_list <- list()
for(x in names(alldat)){
    p_list[[x]] <- DimPlot(alldat[[x]],group.by="celltype",label=TRUE,repel=TRUE)+ggtitle(x)+NoLegend()+scale_color_manual(values=selected_palette)
}

ggarrange(plotlist=p_list,nrow=3,ncol=3)
```

pca
```{r fig.height=7.5, fig.width=8.5}
p_list <- list()
for(x in names(alldat)){
    p_list[[x]] <- DimPlot(alldat[[x]],group.by="celltype",label=TRUE,repel=TRUE,reduction="pca")+ggtitle(x)+NoLegend()+scale_color_manual(values=selected_palette)
}

ggarrange(plotlist=p_list,nrow=3,ncol=3)
```

#### zoom in
```{r fig.height=2.5, fig.width=16.5, message=FALSE, warning=FALSE}
processSubset <- function(x){
    tmp_sub <- subset(x,celltype%in%c("H1975","H1975.zero"))
    tmp_sub <- RunPCA(tmp_sub)
    tmp_sub <- RunUMAP(tmp_sub,dims=1:50)
    return(tmp_sub)
}
alldat_sub <- lapply(alldat,processSubset)

umap_list <- list()
for(x in names(alldat_sub)){
    umap_list[[x]] <- DimPlot(alldat_sub[[x]],group.by="celltype",repel=TRUE)+ggtitle(x)+scale_color_manual(values=selected_palette)
}
cb10x5cl_umap <- ggarrange(plotlist=umap_list,nrow=1,ncol=5)
cb10x5cl_umap

pca_list <- list()
for(x in names(alldat_sub)){
    # get variance explained
    tmppca <- alldat_sub[[x]][["pca"]]
    total_variance <- tmppca@misc$total.variance
    # Get the total/top50 variance
    eigValues <- (tmppca@stdev)^2  ## EigenValues
    varExplained <- round(eigValues/total_variance*100,2)[1:2] #total
    #varExplained <- round(eigValues/sum(eigValues)*100,2)[1:2] #top50
    rm(tmppca)
    
    pca_list[[x]] <- DimPlot(alldat_sub[[x]],group.by="celltype",repel=TRUE,reduction="pca")+ggtitle(x)+scale_color_manual(values=selected_palette)+xlab(paste0("PC1(",varExplained[1],"%)"))+ylab(paste0("PC2(",varExplained[2],"%)"))
}

cb10x5cl_pca <- ggarrange(plotlist=pca_list,nrow=1,ncol=5)
cb10x5cl_pca
```

## GSE75748
```{r}
##bulk
bulk_dat <- read.csv("GSE75748_bulk_cell_type_ec.csv",header=TRUE,row.names=1)
bulk_dat <- DGEList(bulk_dat)
keep <- filterByExpr(bulk_dat)
bulk_dat <- bulk_dat[keep,,keep.lib.sizes=FALSE]
bulk_dat <- calcNormFactors(bulk_dat)
bulk_meta <- c("H1","H1","H1","H1","H9","H9","H9","DEC","DEC","EC","EC","EC","HFF","HFF","HFF","NPC","NPC","TB","TB")
mm <- model.matrix(~0 + bulk_meta)
bulk_dat <- voom(bulk_dat,mm,plot=FALSE)$E
bulk_dat <- data.frame(H1=apply(bulk_dat[,1:4],1,mean),H9=apply(bulk_dat[,5:7],1,mean),DEC=apply(bulk_dat[,8:9],1,mean),EC=apply(bulk_dat[,10:12],1,mean),HFF=apply(bulk_dat[,13:15],1,mean),NPC=apply(bulk_dat[,16:17],1,mean),TB=apply(bulk_dat[,18:19],1,mean))

##raw
dat <- fread("GSE75748_sc_cell_type_genebycell.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
dat <- as.matrix(dat,rownames=1)

metadat <- sapply(colnames(dat),function(x) strsplit(x,"\\_")[[1]][1])
metadat <- data.frame(cell=colnames(dat),celltype=metadat)

bulk_dat <- bulk_dat[rownames(bulk_dat)%in%rownames(dat),]
dat <- dat[rownames(bulk_dat),]

#raw lognorm
dat <- CreateSeuratObject(dat,meta.data=metadat,min.cells=0,min.features=0)
dat <- NormalizeData(dat)
dat <- FindVariableFeatures(dat,nfeatures=3000)
dat <- ScaleData(dat)
dat <- RunPCA(dat)
dat <- RunUMAP(dat,dims=1:10)
```

### randomly assign top gene zeros for each cell
50% H1 and 7400 gene(~60%)
```{r}
#rank expressed features in selected types
SelectedType <- rownames(dat@meta.data[dat@meta.data$celltype=="H1",])
SelectedTypeMat <- as.matrix(dat@assays$RNA@counts)
SelectedTypeMat <- SelectedTypeMat[,SelectedType]
TopFeatures <- apply(SelectedTypeMat,1,sum)
names(TopFeatures) <- rownames(SelectedTypeMat)
TopFeatures <- names(TopFeatures[order(TopFeatures,decreasing=TRUE)])

#select 10% of cells and copy
set.seed(1)
Selected10Cells <- sample(x=SelectedType,size=0.5*length(SelectedType))
SelectedCellMat <- SelectedTypeMat[,Selected10Cells]

#assign zeros
RandomAssignZeros <- function(x,g,gname){
    names(x) <- gname
    SelectedGenes <- sample(g,size=7400)
    x[SelectedGenes] <- 0
    return(x)
}

ProcessedCellMat <- apply(SelectedCellMat,2,RandomAssignZeros,TopFeatures,rownames(SelectedCellMat))
colnames(ProcessedCellMat) <- paste0(colnames(ProcessedCellMat),".copy")

#combine
tmp <- cbind(as.matrix(dat@assays$RNA@counts),ProcessedCellMat)
tmpmeta <- data.frame(cell=colnames(ProcessedCellMat),celltype=rep("H1.zero",ncol(ProcessedCellMat)))
rownames(tmpmeta) <- tmpmeta$cell
tmpmeta <- rbind(dat@meta.data[,c("cell","celltype")],tmpmeta)

#plot
tmp <- CreateSeuratObject(tmp,meta.data=tmpmeta,min.cells=0,min.features=0)

tmp <- NormalizeData(tmp)
tmp <- FindVariableFeatures(tmp,nfeatures=3000)
tmp <- ScaleData(tmp)
tmp <- RunPCA(tmp)
tmp <- RunUMAP(tmp,dims=1:10)

#sct
tmp_sct <- CreateSeuratObject(counts=tmp@assays$RNA@counts,meta.data=tmpmeta,min.cells=0,min.features=0)
tmp_sct <- SCTransform(tmp_sct,vst.flavor="v2",verbose=FALSE,min_cells=0)
tmp_sct <- RunPCA(tmp_sct)
tmp_sct <- RunUMAP(tmp_sct,dims=1:10)
```

CoDA
```{r}
#add pseudocount s/gm to raw count
tmp1 <- as.matrix(tmp@assays$RNA@counts)
gm_s <- exp(mean(log(apply(tmp1,2,sum))))
tmp1 <- apply(tmp1,2,function(x) x+(sum(x)/gm_s) )
tmp1 <- t(CLOSE(t(tmp1)))

#CLR
raw_coda_clr <- apply(tmp1,2,function(x) x/exp(mean(log(x))) )
raw_coda_clr <- log2(raw_coda_clr)

# HKGLR
HKGenes <- which(rownames(tmp1)%in%c("ACTB","ARBP","GAPDH","HPRT","SDHA","UBC","YWHAZ"))
raw_coda_ilr <- apply(tmp1,2,function(x) x/exp(mean(log(x[HKGenes]))) )
raw_coda_ilr <- log2(raw_coda_ilr)

##ILR
cb_ILR <- as.matrix(t(coda.base::coordinates(t(tmp1))))
cb_ILR <- cb_ILR[,colnames(raw_coda_clr)]

#Seurat process
processSeurat <- function(x,metadat){
    res <- CreateSeuratObject(x,meta.data=metadat,min.cells=0,min.features=0)
    res <- FindVariableFeatures(res,nfeatures=3000)
    res <- ScaleData(res)
    res <- RunPCA(res)
    res <- RunUMAP(res,dims=1:50)
    return(res)
}

raw_coda_clr <- processSeurat(raw_coda_clr,tmp@meta.data)
raw_coda_ilr <- processSeurat(raw_coda_ilr,tmp@meta.data)
cb_ILR <- processSeurat(cb_ILR,tmp@meta.data)

alldat <- list(Raw_LogNorm=tmp,Raw_SCT=tmp_sct,SGM_CLR=raw_coda_clr,SGM_ILR=cb_ILR,SGM_HKGLR=raw_coda_ilr)
alldat <- alldat[c("Raw_LogNorm","Raw_SCT","SGM_CLR","SGM_ILR","SGM_HKGLR")]
alldat_GSE75748 <- alldat
```

plots umap
```{r fig.height=7.5, fig.width=8.5}
p_list <- list()
for(x in names(alldat)){
    p_list[[x]] <- DimPlot(alldat[[x]],group.by="celltype",label=TRUE,repel=TRUE)+ggtitle(x)+NoLegend()+scale_color_manual(values=selected_palette)
}

p_list <- p_list[c("LogNorm","SCT","mdCLR","CLR","IQLR","LVHA","HKGLR","IQLR_group","LVHA_group")]
ggarrange(plotlist=p_list,nrow=3,ncol=3)
```

#### zoom in
```{r fig.height=2.5, fig.width=16.5, message=FALSE, warning=FALSE}
processSubset <- function(x){
    tmp_sub <- subset(x,celltype%in%c("H1","H1.zero","H9"))
    tmp_sub <- RunPCA(tmp_sub)
    tmp_sub <- RunUMAP(tmp_sub,dims=1:10)
    return(tmp_sub)
}
alldat_sub <- lapply(alldat,processSubset)

umap_list <- list()
for(x in names(alldat_sub)){
    umap_list[[x]] <- DimPlot(alldat_sub[[x]],group.by="celltype",repel=TRUE)+ggtitle(x)+scale_color_manual(values=selected_palette)
}
gse75748_umap <- ggarrange(plotlist=umap_list,nrow=1,ncol=5)
gse75748_umap

pca_list <- list()
for(x in names(alldat_sub)){
    # get variance explained
    tmppca <- alldat_sub[[x]][["pca"]]
    total_variance <- tmppca@misc$total.variance
    # Get the total/top50 variance
    eigValues <- (tmppca@stdev)^2  ## EigenValues
    varExplained <- round(eigValues/total_variance*100,2)[1:2] #total
    #varExplained <- round(eigValues/sum(eigValues)*100,2)[1:2] #top50
    rm(tmppca)
    
    pca_list[[x]] <- DimPlot(alldat_sub[[x]],group.by="celltype",repel=TRUE,reduction="pca")+ggtitle(x)+scale_color_manual(values=selected_palette)+xlab(paste0("PC1(",varExplained[1],"%)"))+ylab(paste0("PC2(",varExplained[2],"%)"))
}
#ggarrange(plotlist=pca_list,nrow=3,ncol=3)
gse75748_pca <- ggarrange(plotlist=pca_list,nrow=1,ncol=5)
gse75748_pca
```

## combine figure
```{r fig.height=5, fig.width=16.5}
ggarrange(cb10x5cl_umap,gse75748_umap,ncol=1,nrow=2)
ggarrange(cb10x5cl_pca,gse75748_pca,ncol=1,nrow=2)
```

## clustering performance of subset degraded-cell datasets
### GSE75748
```{r}
alldat_GSE75748_sub <- lapply(alldat_GSE75748,function(x) subset(x,celltype%in%c("H1","H1.zero","H9")) )
metadat <- alldat_GSE75748_sub$Raw_LogNorm@meta.data
alldat_GSE75748_sub1 <- lapply(alldat_GSE75748_sub,function(x) as.matrix(x@assays$RNA@counts) )
alldat_GSE75748_sub1$Raw_LogNorm <- as.matrix(alldat_GSE75748_sub$Raw_LogNorm@assays$RNA@data)
alldat_GSE75748_sub1$Raw_SCT <- as.matrix(alldat_GSE75748_sub$Raw_SCT@assays$SCT@data)

datCluster <- lapply(alldat_GSE75748_sub1,getCluster,metadat,cent=3,scale_data=TRUE)

res1_lv <- as.data.frame(lapply(datCluster,evalClustering))
as.data.frame(t(res1_lv))

res1_km <- as.data.frame(lapply(datCluster,evalClustering,method="kmeans"))
as.data.frame(t(res1_km))
```

### CellBench
```{r}
alldat_CellBench_sub <- lapply(alldat_CellBench,function(x) subset(x,celltype%in%c("H1975","H1975.zero")) )
metadat <- alldat_CellBench_sub$Raw_LogNorm@meta.data
alldat_CellBench_sub1 <- lapply(alldat_CellBench_sub,function(x) as.matrix(x@assays$RNA@counts) )
alldat_CellBench_sub1$Raw_LogNorm <- as.matrix(alldat_CellBench_sub$Raw_LogNorm@assays$RNA@data)
alldat_CellBench_sub1$Raw_SCT <- as.matrix(alldat_CellBench_sub$Raw_SCT@assays$SCT@data)

datCluster <- lapply(alldat_CellBench_sub1,getCluster,metadat,cent=2,scale_data=TRUE)

res2_lv <- as.data.frame(lapply(datCluster,evalClustering))
as.data.frame(t(res2_lv))

res2_km <- as.data.frame(lapply(datCluster,evalClustering,method="kmeans"))
as.data.frame(t(res2_km))
```

### summary and visualization
#### process data(lv)
```{r}
##LV cluster
rownames(res1_lv) <- c("Hacc","Hpur","ARI","medianSil","NMI")

res_lv <- data.frame()
res_lv_median <- data.frame()
for(i in 1:2){
    tmp1 <- res1_lv[i,1]-res1_lv[i,-1]
    tmp2 <- res2_lv[i,1]-res2_lv[i,-1]
    tmp <- rbind(tmp1,tmp2)
    rownames(tmp) <- paste0(rownames(res1_lv)[i],"_",c("GSE75748","cellbench"))
    tmp_median <- apply(tmp,2,median,na.rm=TRUE)
    res_lv <- rbind(res_lv,tmp)
    res_lv_median <- rbind(res_lv_median,tmp_median)
}

for(i in 3:5){
    tmp1 <- res1_lv[i,-1]-res1_lv[i,1]
    tmp2 <- res2_lv[i,-1]-res2_lv[i,1]
    tmp <- rbind(tmp1,tmp2)
    rownames(tmp) <- paste0(rownames(res1_lv)[i],"_",c("GSE75748","cellbench"))
    tmp_median <- apply(tmp,2,median,na.rm=TRUE)
    res_lv <- rbind(res_lv,tmp)
    res_lv_median <- rbind(res_lv_median,tmp_median)
}

res_lv_median <- as.data.frame(t(res_lv_median))
rownames(res_lv_median) <- colnames(res1_lv)[-1]
colnames(res_lv_median) <- c("Hacc","Hpur","ARI","medianSil","NMI")

res_lv_median$method <- rownames(res_lv_median)
res_lv_median_melt <- data.table::melt(res_lv_median,id.vars=c("method"))
colnames(res_lv_median_melt)[2:3] <- c("Metrics","Metrics_Difference")
head(res_lv_median_melt)

res_lv$Metrics_Datasets <- rownames(res_lv)
res_lv$Metrics <- sapply(rownames(res_lv),function(x) strsplit(x,"\\_")[[1]][1] )
res_lv_melt <- data.table::melt(res_lv,id.vars=c("Metrics_Datasets","Metrics"))
colnames(res_lv_melt)[3:4] <- c("method","Metrics_Difference")
head(res_lv_melt)
```

heatmap(all datasets&metrics)
```{r}
rank1 <- apply(res_lv_median[,-ncol(res_lv_median)],2,rank)
rank1 <- apply(rank1,1,median)
rank1 <- names(rank1[order(rank1)])
res_lv_melt$method <- factor(res_lv_melt$method,levels=rank1)

tmp_hm <- res_lv_melt[,]
zero_s_lv_hm_ind <- ggplot(tmp_hm,aes(method,Metrics_Datasets,fill=Metrics_Difference))+geom_tile()+theme(axis.text.x=element_text(angle=90))+scale_fill_gradient2(low="blue",mid="white",high="red",midpoint=0)+geom_hline(yintercept=c(2.5,4.5,6.5),linetype="dashed")+ggtitle("Louvain Clustering Evaluation Metrics(Values of Difference)")+geom_text(aes(label=round(Metrics_Difference,2)))+xlab("Method")
zero_s_lv_hm_ind
```

#### process data(km)
```{r}
##LV cluster
rownames(res1_km) <- c("Hacc","Hpur","ARI","medianSil","NMI")

res_km <- data.frame()
res_km_median <- data.frame()
for(i in 1:2){
    tmp1 <- res1_km[i,1]-res1_km[i,-1]
    tmp2 <- res2_km[i,1]-res2_km[i,-1]
    tmp <- rbind(tmp1,tmp2)
    rownames(tmp) <- paste0(rownames(res1_km)[i],"_",c("GSE75748","cellbench"))
    tmp_median <- apply(tmp,2,median,na.rm=TRUE)
    res_km <- rbind(res_km,tmp)
    res_km_median <- rbind(res_km_median,tmp_median)
}

for(i in 3:5){
    tmp1 <- res1_km[i,-1]-res1_km[i,1]
    tmp2 <- res2_km[i,-1]-res2_km[i,1]
    tmp <- rbind(tmp1,tmp2)
    rownames(tmp) <- paste0(rownames(res1_km)[i],"_",c("GSE75748","cellbench"))
    tmp_median <- apply(tmp,2,median,na.rm=TRUE)
    res_km <- rbind(res_km,tmp)
    res_km_median <- rbind(res_km_median,tmp_median)
}

res_km_median <- as.data.frame(t(res_km_median))
rownames(res_km_median) <- colnames(res1_km)[-1]
colnames(res_km_median) <- c("Hacc","Hpur","ARI","medianSil","NMI")

res_km_median$method <- rownames(res_km_median)
res_km_median_melt <- data.table::melt(res_km_median,id.vars=c("method"))
colnames(res_km_median_melt)[2:3] <- c("Metrics","Metrics_Difference")
head(res_km_median_melt)

res_km$Metrics_Datasets <- rownames(res_km)
res_km$Metrics <- sapply(rownames(res_km),function(x) strsplit(x,"\\_")[[1]][1] )
res_km_melt <- data.table::melt(res_km,id.vars=c("Metrics_Datasets","Metrics"))
colnames(res_km_melt)[3:4] <- c("method","Metrics_Difference")
head(res_km_melt)
```

heatmap(all datasets&metrics)
```{r}
rank1 <- apply(res_km_median[,-ncol(res_km_median)],2,rank)
rank1 <- apply(rank1,1,median)
rank1 <- names(rank1[order(rank1)])
res_km_melt$method <- factor(res_km_melt$method,levels=rank1)

tmp_hm <- res_km_melt[,]
#tmp_hm$Metrics_Difference <- ifelse(tmp_hm$Metrics_Difference<(-0.25),-0.25,tmp_hm$Metrics_Difference)
#tmp_hm$Metrics_Difference <- ifelse(tmp_hm$Metrics_Difference>0.1,0.1,tmp_hm$Metrics_Difference)
#tmp_hm <- tmp_hm[tmp_hm$Metrics!="medianSil",]
zero_s_km_hm_ind <- ggplot(tmp_hm,aes(method,Metrics_Datasets,fill=Metrics_Difference))+geom_tile()+theme(axis.text.x=element_text(angle=90))+scale_fill_gradient2(low="blue",mid="white",high="red",midpoint=0)+geom_hline(yintercept=c(2.5,4.5,6.5),linetype="dashed")+ggtitle("K-means Clustering Evaluation Metrics(Values of Difference)")+geom_text(aes(label=round(Metrics_Difference,2)))+xlab("Method")
zero_s_km_hm_ind
```

### combine
```{r fig.height=1.8, fig.width=6.6}
ggarrange(zero_s_lv_hm_ind,zero_s_km_hm_ind,nrow=1,ncol=2)
```

