---
title: "CoDA_Scalability"
output: html_document
---

library
```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(Seurat)
library(scran)
library(SeuratData)
library(edgeR)
library(limma)
library(clusterProfiler)
library(data.table)
library(SingleCellExperiment)
library(SummarizedExperiment)
library(muscat)
library(pROC)
library(igraph)
library(mclust)
library(cluster)
library(SCINA)
library(caret)
library(robCompositions)
library(easyCODA)
library(MAST)
library(AUCell)
library(HGNChelper)
library(SeuratDisk)
library(scmap)
library(monocle3)
library(ggpubr)
library(slingshot)
library(grDevices)
library(RColorBrewer)
library(randomForest)
library(caTools)
library(mltools)
library(stringr)
library(reldist)
library(aricode)
library(TSCAN)
library(reshape2)
library(pheatmap)
library(philentropy)
library(rawr)
library(splatter)
library(MatrixCorrelation)
library(scater)
library(Hmisc)
library(peakRAM)
```

clustering functions
```{r message=FALSE, warning=FALSE}
getCluster <- function(x,meta,reso=0.01,npc=10,cent=4,scale_data=FALSE){
    tmp <- CreateSeuratObject(counts=x,meta.data=meta,min.cells=0,min.features=0)
    tmp <- FindVariableFeatures(tmp,nfeatures=3000,verbose=FALSE)
    if(scale_data){
        tmp <- ScaleData(tmp,verbose=FALSE)
    }else{
        tmp[["RNA"]]@scale.data <- as.matrix(tmp[["RNA"]]@counts[VariableFeatures(tmp),])
    }
    tmp <- RunPCA(tmp,verbose=FALSE)
    tmp$kmeans <- kmeans(tmp@reductions$pca@cell.embeddings[,1:npc],centers=cent)$cluster
    tmp <- FindNeighbors(tmp,dims=1:npc,verbose=FALSE)
    tmp <- FindClusters(tmp,resolution=reso,verbose=FALSE)
    tmpgraph <- buildSNNGraph(tmp@reductions$pca@cell.embeddings[,1:npc],transposed=T,k=10,d=NA)
    res <- cluster_louvain(tmpgraph)$membership
    cc <- aggregate(tmp@reductions$pca@cell.embeddings[,1:npc],list(res),mean)
    cc <- as.matrix(cc[,-1])
    hclu <- hclust(dist(cc))
    clu <- cutree(hclu,cent)
    clu <- clu[res]      
    tmp$lv_clu <- clu
    #tmp <- tmp@meta.data
    tmp <- RunUMAP(tmp,dims=1:npc,verbose=FALSE)
    return(tmp)
}

evalClustering <- function(x,npc=10,method="lv_clu"){
    
    acc <- -mean(sapply(unique(x@meta.data[,method]),function(i){
        p <- table(x@meta.data[x@meta.data[,method]==i,"celltype"])/sum(x@meta.data[,method]==i)
        sum(p*log(p),na.rm=TRUE)
    }))

    pur <- -mean(sapply(unique(x$celltype),function(sct){
        p <- table(x@meta.data[x$celltype==sct,method])/sum(x$celltype==sct)
        sum(p*log(p),na.rm=TRUE)
    }))

    ARI <- adjustedRandIndex(x$celltype,x@meta.data[,method])

    d <- dist(x@reductions$pca@cell.embeddings[,1:npc])
    s <- silhouette(x@meta.data[,method],dist=d)
    medianSil <- median(s[,3])
    
    NMI_value <- NMI(x@meta.data[,method],x$celltype)
    #NMI_value <- NMI(x$lv_clu,x$celltype,variant="sum")
    
    df <- c(acc,pur,ARI,medianSil,NMI_value)
    names(df) <- c("acc","pur","ARI","medianSil","NMI")
    return(df)
}
```

# 1500
read data
```{r message=FALSE, warning=FALSE}
##raw
raw_dat <- fread("Sim_CB10x5cl_G10kC1500_5UGroup_dropout92.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
raw_dat <- as.matrix(raw_dat,rownames=1)

metadat <- readRDS("Sim_CB10x5cl_G10kC1500_5UGroup_dropout92.rds")
truedat <- metadat@assays@data$TrueCounts

metadat <- as.data.frame(colData(metadat))
metadat$cell <- metadat$Cell
metadat$celltype <- metadat$Group
```

## CoDA
```{r}
start.time <- Sys.time()

mem <- peakRAM({
  #add pseudocount s/gm to raw count
  tmp <- as.matrix(raw_dat)
  gm_s <- exp(mean(log(apply(tmp,2,sum))))
  tmp <- apply(tmp,2,function(x) x+(sum(x)/gm_s) )
  tmp <- t(CLOSE(t(tmp)))

  #CLR
  raw_coda_clr <- apply(tmp,2,function(x) x/exp(mean(log(x))) )
  raw_coda_clr <- log2(raw_coda_clr)
})

end.time <- Sys.time()

time_1500_CLRtrans <- end.time-start.time
time_1500_CLRtrans
mem_1500_CLRtrans <- mem$Peak_RAM_Used_MiB
mem_1500_CLRtrans
```


## normalize
normalized data
```{r}
raw_lognorm <- CreateSeuratObject(counts=raw_dat,meta.data=metadat,min.cells=0,min.features=0)
start.time <- Sys.time()
mem <- peakRAM({raw_lognorm <- NormalizeData(raw_lognorm)})
end.time <- Sys.time()
time_1500_LogNormtrans <- end.time-start.time
time_1500_LogNormtrans
mem_1500_LogNormtrans <- mem$Peak_RAM_Used_MiB
mem_1500_LogNormtrans
raw_lognorm <- as.matrix(raw_lognorm@assays$RNA@data)

raw_sct <- CreateSeuratObject(counts=raw_dat,meta.data=metadat,min.cells=0,min.features=0)
start.time <- Sys.time()
mem <- peakRAM({raw_sct <- SCTransform(raw_sct,vst.flavor="v2",verbose=FALSE,min_cells=0)})
end.time <- Sys.time()
time_1500_SCTtrans <- end.time-start.time
time_1500_SCTtrans
mem_1500_SCTtrans <- mem$Peak_RAM_Used_MiB
mem_1500_SCTtrans
raw_sct <- as.matrix(raw_sct@assays$SCT@data)
```

writeout/read
```{r}
#writeout
trans_time_res <- data.frame(Method=c("LogNorm","SCT","CLR"),Cell=c("1500","1500","1500"),Transformation_time=as.numeric(c(time_1500_LogNormtrans,time_1500_SCTtrans,time_1500_CLRtrans)))
trans_mem_res <- data.frame(Method=c("LogNorm","SCT","CLR"),Cell=c("1500","1500","1500"),Transformation_mem=c(mem_1500_LogNormtrans,mem_1500_SCTtrans,mem_1500_CLRtrans))

all_res <- list(trans_time_res=trans_time_res,trans_mem_res=trans_mem_res)
saveRDS(all_res,"results/Scalability/Cell1500.rds")
```

# 5000
cell level
read data
```{r message=FALSE, warning=FALSE}
##raw
raw_dat <- fread("ForScal_Sim_Mock_G10kC5000_5UGroup_de0.5_dropout90.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
raw_dat <- as.matrix(raw_dat,rownames=1)
```

## CoDA
```{r}
start.time <- Sys.time()

mem <- peakRAM({
  #add pseudocount s/gm to raw count
  tmp <- as.matrix(raw_dat)
  gm_s <- exp(mean(log(apply(tmp,2,sum))))
  tmp <- apply(tmp,2,function(x) x+(sum(x)/gm_s) )
  tmp <- t(CLOSE(t(tmp)))

  #CLR
  raw_coda_clr <- apply(tmp,2,function(x) x/exp(mean(log(x))) )
  raw_coda_clr <- log2(raw_coda_clr)
})

end.time <- Sys.time()

time_5k_CLRtrans <- end.time-start.time
time_5k_CLRtrans
mem_5k_CLRtrans <- mem$Peak_RAM_Used_MiB
mem_5k_CLRtrans
```


## normalize
normalized data
```{r}
raw_lognorm <- CreateSeuratObject(counts=raw_dat,min.cells=0,min.features=0)
start.time <- Sys.time()
mem <- peakRAM({raw_lognorm <- NormalizeData(raw_lognorm)})
end.time <- Sys.time()
time_5k_LogNormtrans <- end.time-start.time
time_5k_LogNormtrans
mem_5k_LogNormtrans <- mem$Peak_RAM_Used_MiB
mem_5k_LogNormtrans
raw_lognorm <- as.matrix(raw_lognorm@assays$RNA@data)

raw_sct <- CreateSeuratObject(counts=raw_dat,min.cells=0,min.features=0)
start.time <- Sys.time()
mem <- peakRAM({raw_sct <- SCTransform(raw_sct,vst.flavor="v2",verbose=FALSE,min_cells=0)})
end.time <- Sys.time()
time_5k_SCTtrans <- end.time-start.time
time_5k_SCTtrans
mem_5k_SCTtrans <- mem$Peak_RAM_Used_MiB
mem_5k_SCTtrans
raw_sct <- as.matrix(raw_sct@assays$SCT@data)
```

writeout/read
```{r}
#writeout
trans_time_res <- data.frame(Method=c("LogNorm","SCT","CLR"),Cell=c("5000","5000","5000"),Transformation_time=as.numeric(c(time_5k_LogNormtrans,time_5k_SCTtrans,time_5k_CLRtrans)))
trans_mem_res <- data.frame(Method=c("LogNorm","SCT","CLR"),Cell=c("5000","5000","5000"),Transformation_mem=c(mem_5k_LogNormtrans,mem_5k_SCTtrans,mem_5k_CLRtrans))

all_res <- list(trans_time_res=trans_time_res,trans_mem_res=trans_mem_res)
saveRDS(all_res,"results/Scalability/Cell5k.rds")
```

# 10000
read data
```{r message=FALSE, warning=FALSE}
##raw
raw_dat <- fread("ForScal_Sim_Mock_G10kC10k_5UGroup_de0.5_dropout90.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
raw_dat <- as.matrix(raw_dat,rownames=1)
```

## CoDA
```{r}
start.time <- Sys.time()

mem <- peakRAM({
  #add pseudocount s/gm to raw count
  tmp <- as.matrix(raw_dat)
  gm_s <- exp(mean(log(apply(tmp,2,sum))))
  tmp <- apply(tmp,2,function(x) x+(sum(x)/gm_s) )
  tmp <- t(CLOSE(t(tmp)))

  #CLR
  raw_coda_clr <- apply(tmp,2,function(x) x/exp(mean(log(x))) )
  raw_coda_clr <- log2(raw_coda_clr)
})

end.time <- Sys.time()

time_10k_CLRtrans <- end.time-start.time
time_10k_CLRtrans
mem_10k_CLRtrans <- mem$Peak_RAM_Used_MiB
mem_10k_CLRtrans
```

## normalize
normalized data
```{r}
raw_lognorm <- CreateSeuratObject(counts=raw_dat,min.cells=0,min.features=0)
start.time <- Sys.time()
mem <- peakRAM({raw_lognorm <- NormalizeData(raw_lognorm)})
end.time <- Sys.time()
time_10k_LogNormtrans <- end.time-start.time
time_10k_LogNormtrans
mem_10k_LogNormtrans <- mem$Peak_RAM_Used_MiB
mem_10k_LogNormtrans
raw_lognorm <- as.matrix(raw_lognorm@assays$RNA@data)

raw_sct <- CreateSeuratObject(counts=raw_dat,min.cells=0,min.features=0)
start.time <- Sys.time()
mem <- peakRAM({raw_sct <- SCTransform(raw_sct,vst.flavor="v2",verbose=FALSE,min_cells=0)})
end.time <- Sys.time()
time_10k_SCTtrans <- end.time-start.time
time_10k_SCTtrans
mem_10k_SCTtrans <- mem$Peak_RAM_Used_MiB
mem_10k_SCTtrans
raw_sct <- as.matrix(raw_sct@assays$SCT@data)
```

writeout/read
```{r}
#writeout
trans_time_res <- data.frame(Method=c("LogNorm","SCT","CLR"),Cell=c("10000","10000","10000"),Transformation_time=as.numeric(c(time_10k_LogNormtrans,time_10k_SCTtrans,time_10k_CLRtrans)))
trans_mem_res <- data.frame(Method=c("LogNorm","SCT","CLR"),Cell=c("10000","10000","10000"),Transformation_mem=c(mem_10k_LogNormtrans,mem_10k_SCTtrans,mem_10k_CLRtrans))

all_res <- list(trans_time_res=trans_time_res,trans_mem_res=trans_mem_res)
saveRDS(all_res,"results/Scalability/Cell10k.rds")
```

# 50000
read data
```{r message=FALSE, warning=FALSE}
##raw
raw_dat <- fread("ForScal_Sim_Mock_G10kC50k_5UGroup_de0.5_dropout90.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
raw_dat <- as.matrix(raw_dat,rownames=1)
```

## CoDA
```{r}
start.time <- Sys.time()

mem <- peakRAM({
  #add pseudocount s/gm to raw count
  tmp <- as.matrix(raw_dat)
  gm_s <- exp(mean(log(apply(tmp,2,sum))))
  tmp <- apply(tmp,2,function(x) x+(sum(x)/gm_s) )
  tmp <- t(CLOSE(t(tmp)))

  #CLR
  raw_coda_clr <- apply(tmp,2,function(x) x/exp(mean(log(x))) )
  raw_coda_clr <- log2(raw_coda_clr)
})

end.time <- Sys.time()

time_50k_CLRtrans <- end.time-start.time
time_50k_CLRtrans
mem_50k_CLRtrans <- mem$Peak_RAM_Used_MiB
mem_50k_CLRtrans
```

## normalize
normalized data
```{r}
raw_lognorm <- CreateSeuratObject(counts=raw_dat,min.cells=0,min.features=0)
start.time <- Sys.time()
mem <- peakRAM({raw_lognorm <- NormalizeData(raw_lognorm)})
end.time <- Sys.time()
time_50k_LogNormtrans <- end.time-start.time
time_50k_LogNormtrans
mem_50k_LogNormtrans <- mem$Peak_RAM_Used_MiB
mem_50k_LogNormtrans
raw_lognorm <- as.matrix(raw_lognorm@assays$RNA@data)

raw_sct <- CreateSeuratObject(counts=raw_dat,min.cells=0,min.features=0)
start.time <- Sys.time()
mem <- peakRAM({raw_sct <- SCTransform(raw_sct,vst.flavor="v2",verbose=FALSE,min_cells=0)})
end.time <- Sys.time()
time_50k_SCTtrans <- end.time-start.time
time_50k_SCTtrans
mem_50k_SCTtrans <- mem$Peak_RAM_Used_MiB
mem_50k_SCTtrans
raw_sct <- as.matrix(raw_sct@assays$SCT@data)
```

writeout/read
```{r}
#writeout
trans_time_res <- data.frame(Method=c("LogNorm","SCT","CLR"),Cell=c("50000","50000","50000"),Transformation_time=as.numeric(c(time_50k_LogNormtrans,time_50k_SCTtrans,time_50k_CLRtrans)))
trans_mem_res <- data.frame(Method=c("LogNorm","SCT","CLR"),Cell=c("50000","50000","50000"),Transformation_mem=c(mem_50k_LogNormtrans,mem_50k_SCTtrans,mem_50k_CLRtrans))

all_res <- list(trans_time_res=trans_time_res,PCA_time_res=PCA_time_res,trans_mem_res=trans_mem_res,PCA_mem_res=PCA_mem_res)
saveRDS(all_res,"results/Scalability/Cell50k.rds")
```

# summary
```{r}
d1 <- readRDS("results/Scalability/Cell1500.rds")
d2 <- readRDS("results/Scalability/Cell5k.rds")
d3 <- readRDS("results/Scalability/Cell10k.rds")
d4 <- readRDS("results/Scalability/Cell50k.rds")
```

## Transformation time
```{r}
trans_time <- rbind(d1$trans_time_res,d2$trans_time_res,d3$trans_time_res,d4$trans_time_res)

trans_time$Cell <- factor(trans_time$Cell,levels=c("1500","5000","10000","50000"))
```

plots
```{r}
p1_trans_time <- ggplot(trans_time,aes(x=Cell,y=Transformation_time,group=Method,color=Method))+geom_line()+geom_point()+theme_minimal()+xlab("Number of Cells")+ylab("Transformation time(s) used")+ggtitle("Transformation time(s) used")+scale_color_manual(values=selected_palette)+theme(plot.title=element_text(size=18),legend.title=element_text(size=17),legend.text=element_text(size=16),axis.title=element_text(size=17),axis.text=element_text(size=16))
p1_trans_time
```

## Transformation mem
```{r}
trans_mem <- rbind(d1$trans_mem_res,d2$trans_mem_res,d3$trans_mem_res,d4$trans_mem_res)

trans_mem$Cell <- factor(trans_mem$Cell,levels=c("1500","5000","10000","50000"))
```

plots
```{r}
p3_trans_mem <- ggplot(trans_mem,aes(x=Cell,y=Transformation_mem,group=Method,color=Method))+geom_line()+geom_point()+theme_minimal()+xlab("Number of Cells")+ylab("Transformation max memory usage (MB)")+ggtitle("Transformation max memory usage (MB)")+scale_color_manual(values=selected_palette)+theme(plot.title=element_text(size=18),legend.title=element_text(size=17),legend.text=element_text(size=16),axis.title=element_text(size=17),axis.text=element_text(size=16))
p3_trans_mem
```

## combine plot
```{r fig.height=2.5, fig.width=7}
trans_time_mem <- ggarrange(p1_trans_time,p3_trans_mem,ncol=2,nrow=1,labels="AUTO")
trans_time_mem
```

# LRA vs CLR+Seurat PCA
## 1500
### CLR+Seurat PCA processes(CLR+Scale+RunPCA)
read data
```{r message=FALSE, warning=FALSE}
##raw
raw_dat <- fread("Sim_CB10x5cl_G10kC1500_5UGroup_dropout92.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
raw_dat <- as.matrix(raw_dat,rownames=1)

metadat <- readRDS("Sim_CB10x5cl_G10kC1500_5UGroup_dropout92.rds")
truedat <- metadat@assays@data$TrueCounts

metadat <- as.data.frame(colData(metadat))
metadat$cell <- metadat$Cell
metadat$celltype <- metadat$Group

dat <- CreateSeuratObject(counts=raw_dat,meta.data=metadat,min.cells=0,min.features=0)
dat <- FindVariableFeatures(dat,nfeatures=3000,verbose=FALSE)
selected_features <- dat@assays$RNA@var.features
```

CLRPCA
```{r}
start.time <- Sys.time()

mem <- peakRAM({
    CLRPCA <- as.matrix(raw_dat[selected_features,])
    gm_s <- exp(mean(log(apply(CLRPCA,2,sum))))
    CLRPCA <- apply(CLRPCA,2,function(x) x+(sum(x)/gm_s) )
    CLRPCA <- t(CLOSE(t(CLRPCA)))
    CLRPCA <- apply(CLRPCA,2,function(x) x/exp(mean(log(x))) )
    CLRPCA <- log(CLRPCA)
    CLRPCA <- CreateSeuratObject(CLRPCA,min.cells=0,min.features=0)
    CLRPCA <- ScaleData(CLRPCA,features=rownames(CLRPCA),verbose=FALSE)
    CLRPCA <- RunPCA(CLRPCA,features=rownames(CLRPCA),verbose=FALSE)
})

end.time <- Sys.time()

time_1500_CLRPCA <- end.time-start.time
time_1500_CLRPCA
mem_1500_CLRPCA <- mem$Peak_RAM_Used_MiB
mem_1500_CLRPCA
```

LRA
```{r}
#LRA
start.time <- Sys.time()

mem <- peakRAM({
  tmp_forLRA <- as.matrix(raw_dat[selected_features,])
  gm_s <- exp(mean(log(apply(tmp_forLRA,2,sum))))
  tmp_forLRA <- apply(tmp_forLRA,2,function(x) x+(sum(x)/gm_s) )
  tmp_forLRA <- CLOSE(t(tmp_forLRA))
  tmp3.uwLRA <- LRA(tmp_forLRA,weight=FALSE)
})

end.time <- Sys.time()

time_1500_LRA <- end.time-start.time
time_1500_LRA
mem_1500_LRA <- mem$Peak_RAM_Used_MiB
mem_1500_LRA
```

writeout/read
```{r}
#writeout
PCA_time_res <- data.frame(Method=c("CLR+Seurat PCA","LRA"),Cell=c("1500","1500"),PCA_time=as.numeric(c(time_1500_CLRPCA,time_1500_LRA)))
PCA_mem_res <- data.frame(Method=c("CLR+Seurat PCA","LRA"),Cell=c("1500","1500"),PCA_mem=c(mem_1500_CLRPCA,mem_1500_LRA))

all_res <- list(PCA_time_res=PCA_time_res,PCA_mem_res=PCA_mem_res)
saveRDS(all_res,"results/Scalability/Cell1500_CLRPCA_LRA.rds")
```

## 5000
read data
```{r message=FALSE, warning=FALSE}
##raw
raw_dat <- fread("ForScal_Sim_Mock_G10kC5000_5UGroup_de0.5_dropout90.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
raw_dat <- as.matrix(raw_dat,rownames=1)

dat <- CreateSeuratObject(counts=raw_dat,min.cells=0,min.features=0)
dat <- FindVariableFeatures(dat,nfeatures=3000,verbose=FALSE)
selected_features <- dat@assays$RNA@var.features
```

CLRPCA
```{r}
start.time <- Sys.time()

mem <- peakRAM({
    CLRPCA <- as.matrix(raw_dat[selected_features,])
    gm_s <- exp(mean(log(apply(CLRPCA,2,sum))))
    CLRPCA <- apply(CLRPCA,2,function(x) x+(sum(x)/gm_s) )
    CLRPCA <- t(CLOSE(t(CLRPCA)))
    CLRPCA <- apply(CLRPCA,2,function(x) x/exp(mean(log(x))) )
    CLRPCA <- log(CLRPCA)
    CLRPCA <- CreateSeuratObject(CLRPCA,min.cells=0,min.features=0)
    CLRPCA <- ScaleData(CLRPCA,features=rownames(CLRPCA),verbose=FALSE)
    CLRPCA <- RunPCA(CLRPCA,features=rownames(CLRPCA),verbose=FALSE)
})

end.time <- Sys.time()

time_5000_CLRPCA <- end.time-start.time
time_5000_CLRPCA
mem_5000_CLRPCA <- mem$Peak_RAM_Used_MiB
mem_5000_CLRPCA
```

LRA
```{r}
#LRA
start.time <- Sys.time()

mem <- peakRAM({
  tmp_forLRA <- as.matrix(raw_dat[selected_features,])
  gm_s <- exp(mean(log(apply(tmp_forLRA,2,sum))))
  tmp_forLRA <- apply(tmp_forLRA,2,function(x) x+(sum(x)/gm_s) )
  tmp_forLRA <- CLOSE(t(tmp_forLRA))
  tmp3.uwLRA <- LRA(tmp_forLRA,weight=FALSE)
})

end.time <- Sys.time()

time_5000_LRA <- end.time-start.time
time_5000_LRA
mem_5000_LRA <- mem$Peak_RAM_Used_MiB
mem_5000_LRA
```

writeout/read
```{r}
#writeout
PCA_time_res <- data.frame(Method=c("CLR+Seurat PCA","LRA"),Cell=c("5000","5000"),PCA_time=as.numeric(c(time_5000_CLRPCA,time_5000_LRA)))
PCA_mem_res <- data.frame(Method=c("CLR+Seurat PCA","LRA"),Cell=c("5000","5000"),PCA_mem=c(mem_5000_CLRPCA,mem_5000_LRA))

all_res <- list(PCA_time_res=PCA_time_res,PCA_mem_res=PCA_mem_res)
saveRDS(all_res,"results/Scalability/Cell5000_CLRPCA_LRA.rds")
```

## 10000
read data
```{r message=FALSE, warning=FALSE}
##raw
raw_dat <- fread("ForScal_Sim_Mock_G10kC10k_5UGroup_de0.5_dropout90.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
raw_dat <- as.matrix(raw_dat,rownames=1)

dat <- CreateSeuratObject(counts=raw_dat,min.cells=0,min.features=0)
dat <- FindVariableFeatures(dat,nfeatures=3000,verbose=FALSE)
selected_features <- dat@assays$RNA@var.features
```

CLRPCA
```{r}
start.time <- Sys.time()

mem <- peakRAM({
    CLRPCA <- as.matrix(raw_dat[selected_features,])
    gm_s <- exp(mean(log(apply(CLRPCA,2,sum))))
    CLRPCA <- apply(CLRPCA,2,function(x) x+(sum(x)/gm_s) )
    CLRPCA <- t(CLOSE(t(CLRPCA)))
    CLRPCA <- apply(CLRPCA,2,function(x) x/exp(mean(log(x))) )
    CLRPCA <- log(CLRPCA)
    CLRPCA <- CreateSeuratObject(CLRPCA,min.cells=0,min.features=0)
    CLRPCA <- ScaleData(CLRPCA,features=rownames(CLRPCA),verbose=FALSE)
    CLRPCA <- RunPCA(CLRPCA,features=rownames(CLRPCA),verbose=FALSE)
})

end.time <- Sys.time()

time_10k_CLRPCA <- end.time-start.time
time_10k_CLRPCA
mem_10k_CLRPCA <- mem$Peak_RAM_Used_MiB
mem_10k_CLRPCA
```


LRA
```{r}
#LRA
start.time <- Sys.time()

mem <- peakRAM({
  tmp_forLRA <- as.matrix(raw_dat[selected_features,])
  gm_s <- exp(mean(log(apply(tmp_forLRA,2,sum))))
  tmp_forLRA <- apply(tmp_forLRA,2,function(x) x+(sum(x)/gm_s) )
  tmp_forLRA <- CLOSE(t(tmp_forLRA))
  tmp3.uwLRA <- LRA(tmp_forLRA,weight=FALSE)
})

end.time <- Sys.time()

time_10k_LRA <- end.time-start.time
time_10k_LRA
mem_10k_LRA <- mem$Peak_RAM_Used_MiB
mem_10k_LRA
```

writeout/read
```{r}
#writeout
PCA_time_res <- data.frame(Method=c("CLR+Seurat PCA","LRA"),Cell=c("10000","10000"),PCA_time=as.numeric(c(time_10k_CLRPCA,time_10k_LRA)))
PCA_mem_res <- data.frame(Method=c("CLR+Seurat PCA","LRA"),Cell=c("10000","10000"),PCA_mem=c(mem_10k_CLRPCA,mem_10k_LRA))

all_res <- list(PCA_time_res=PCA_time_res,PCA_mem_res=PCA_mem_res)
saveRDS(all_res,"results/Scalability/Cell10k_CLRPCA_LRA.rds")
```

## 50000
read data
```{r message=FALSE, warning=FALSE}
##raw
raw_dat <- fread("ForScal_Sim_Mock_G10kC50k_5UGroup_de0.5_dropout90.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
raw_dat <- as.matrix(raw_dat,rownames=1)

dat <- CreateSeuratObject(counts=raw_dat,min.cells=0,min.features=0)
dat <- FindVariableFeatures(dat,nfeatures=3000,verbose=FALSE)
selected_features <- dat@assays$RNA@var.features
```

CLRPCA
```{r}
start.time <- Sys.time()

mem <- peakRAM({
    CLRPCA <- as.matrix(raw_dat[selected_features,])
    gm_s <- exp(mean(log(apply(CLRPCA,2,sum))))
    CLRPCA <- apply(CLRPCA,2,function(x) x+(sum(x)/gm_s) )
    CLRPCA <- t(CLOSE(t(CLRPCA)))
    CLRPCA <- apply(CLRPCA,2,function(x) x/exp(mean(log(x))) )
    CLRPCA <- log(CLRPCA)
    CLRPCA <- CreateSeuratObject(CLRPCA,min.cells=0,min.features=0)
    CLRPCA <- ScaleData(CLRPCA,features=rownames(CLRPCA),verbose=FALSE)
    CLRPCA <- RunPCA(CLRPCA,features=rownames(CLRPCA),verbose=FALSE)
})

end.time <- Sys.time()

time_50k_CLRPCA <- end.time-start.time
time_50k_CLRPCA
mem_50k_CLRPCA <- mem$Peak_RAM_Used_MiB
mem_50k_CLRPCA
```


LRA
```{r}
#LRA
start.time <- Sys.time()

mem <- peakRAM({
  tmp_forLRA <- as.matrix(raw_dat[selected_features,])
  gm_s <- exp(mean(log(apply(tmp_forLRA,2,sum))))
  tmp_forLRA <- apply(tmp_forLRA,2,function(x) x+(sum(x)/gm_s) )
  tmp_forLRA <- CLOSE(t(tmp_forLRA))
  tmp3.uwLRA <- LRA(tmp_forLRA,weight=FALSE)
})

end.time <- Sys.time()

time_50k_LRA <- end.time-start.time
time_50k_LRA
mem_50k_LRA <- mem$Peak_RAM_Used_MiB
mem_50k_LRA
```

writeout/read
```{r}
#writeout
PCA_time_res <- data.frame(Method=c("CLR+Seurat PCA","LRA"),Cell=c("50000","50000"),PCA_time=as.numeric(c(time_50k_CLRPCA,time_50k_LRA)))
PCA_mem_res <- data.frame(Method=c("CLR+Seurat PCA","LRA"),Cell=c("50000","50000"),PCA_mem=c(mem_50k_CLRPCA,mem_50k_LRA))

all_res <- list(PCA_time_res=PCA_time_res,PCA_mem_res=PCA_mem_res)
saveRDS(all_res,"results/Scalability/Cell50k_CLRPCA_LRA.rds")
```

## summary
```{r}
d1 <- readRDS("results/Scalability/Cell1500_CLRPCA_LRA.rds")
d2 <- readRDS("results/Scalability/Cell5000_CLRPCA_LRA.rds")
d3 <- readRDS("results/Scalability/Cell10k_CLRPCA_LRA.rds")
d4 <- readRDS("results/Scalability/Cell50k_CLRPCA_LRA.rds")

d3$PCA_time_res$PCA_time <- d3$PCA_time_res$PCA_time*60
```

### PCA time
```{r}
pca_time <- rbind(d1$PCA_time_res,d2$PCA_time_res,d3$PCA_time_res,d4$PCA_time_res)

pca_time$Cell <- factor(pca_time$Cell,levels=c("1500","5000","10000","50000"))
#df_time$Time_log <- log2(df_time$Time)
pca_time$Time_min <- pca_time$PCA_time/60

pca_time[pca_time$Method=="CLR+Seurat PCA",]$Method <- "CLR+partial-SVD"
```

plots
```{r}
p2_pca_time <- ggplot(pca_time,aes(x=Cell,y=Time_min,group=Method,color=Method))+geom_line()+geom_point()+theme_minimal()+xlab("Number of Cells")+ylab("Time(min) used")+ggtitle("Time(min) used (Top 3000 features used)")+scale_color_manual(values=selected_palette)+theme(plot.title=element_text(size=18),legend.title=element_text(size=17),legend.text=element_text(size=16),axis.title=element_text(size=17),axis.text=element_text(size=16))
p2_pca_time
```

### PCA mem
```{r}
pca_mem <- rbind(d1$PCA_mem_res,d2$PCA_mem_res,d3$PCA_mem_res,d4$PCA_mem_res)

pca_mem$Cell <- factor(pca_mem$Cell,levels=c("1500","5000","10000","50000"))

pca_mem[pca_mem$Method=="CLR+Seurat PCA",]$Method <- "CLR+partial-SVD"
```

plots
```{r}
p4_pca_mem <- ggplot(pca_mem,aes(x=Cell,y=PCA_mem,group=Method,color=Method))+geom_line()+geom_point()+theme_minimal()+xlab("Number of Cells")+ylab("Max memory usage (MB)")+ggtitle("Max memory usage (MB) (Top 3000 features used)")+scale_color_manual(values=selected_palette)+theme(plot.title=element_text(size=18),legend.title=element_text(size=17),legend.text=element_text(size=16),axis.title=element_text(size=17),axis.text=element_text(size=16))
p4_pca_mem
```

## combine plot
```{r fig.height=2.5, fig.width=7.5}
pca_time_mem <- ggarrange(p2_pca_time,p4_pca_mem,ncol=2,nrow=1,labels=c("C","D"))
pca_time_mem
```

```{r fig.height=5, fig.width=7.5}
ggarrange(trans_time_mem,pca_time_mem,ncol=1,nrow=2)
```

